\subsection{Target protein: BRD4}

Bromodomain-containing proteins, and in particular the four members of the BET
(bromodomain and extra terminal domain) family, are currently a major focus of
research in the pharmaceutical industry. Small molecule inhibitors of these
proteins have shown promising preclinical efficacy in pathologies ranging from
cancer to inflammation. Indeed, several compounds are progressing through
early stage clinical trials and are showing exciting early
results~\cite{Theodoulou2016}. One of the most extensively studied targets in
this family is the first bromodomain of bromodomain-containing protein 4
(BRD4-BD1) for which extensive crystallographic and ligand binding data are
available~\cite{Bamborough2012}.

We have previously investigated a congeneric series of ligands binding to
BRD4-BD1 (we shall from now on refer to this are simply BRD4) using both
ESMACS and TIES. This was performed in the context of a blind test of the
protocols in collaboration with GlaxoSmithKline~\cite{Wan2017brd4}. The goal
was to benchmark the ESMACS and TIES protocols in a realistic drug discovery
scenario. In the original study, we investigated chemical structures of 16
ligands based on a single tetrahydroquinoline (THQ) scaffold.
% ~\cite{Gosmini2014}. 
Here we focus on the first seven of these ligands to test
and refine the protocols used and the way in which they were executed. The
results of our previous work provide a benchmark of both accuracy and
statistical uncertainty to which we can compare our new results.

Initial coordinates for the protein-ligand system where taken from the X-ray
crystal structure PDB ID: 4BJX.
% ~\cite{Wyce2013}. 
This structure contains a
ligand based on the THQ template and other ligands were alligned with this
common scaffold. Preparation and setup of the simulations were implemented
using our automated tool, BAC~\cite{Sadiq2008}. This process including
parametrization of the compounds, solvation of the complexes, electrostatic
neutralization of the systems by adding counterions and generation of
configurations files for the simulations. The AMBER ff99SB-ILDN
% ~\cite{Lindorff-Larsen2010} 
force field was used for the protein, and TIP3P was used
for water molecules. Compound parameters were produced using the general AMBER
force field (GAFF)~\cite{Wang2004} with Gaussian 03
%~\cite{Frisch} 
to optimize compound geometries and to determine electrostatic potentials at
the Hartree–Fock level with 6-31G** basis functions. The restrained
electrostatic potential (RESP) module in the AMBER package
%~\cite{Case2005} 
was used to calculate the partial atomic charges for the compounds. All systems
were solvated in orthorhombic water boxes with a minimum extension from the
protein of 14 \AA\ (the resulting systems contain approximately 40 thousand
atoms).
