\RequirePackage{amsmath}
\documentclass{llncs}

\input{head}
\input{include}



\title{Rapid, concurrent and adaptive extreme scale binding free energy calculation}

%\title{Concurrent Drug Binding Affinity Computations using Adaptive Simulations}

% \author{
%   \IEEEauthorblockN{Jumana Dakka, Vivek Balasubramanian, Matteo Turilli, Shantenu Jha}
%   \IEEEauthorblockA{RADICAL Laboratory, Electric and Computer Engineering,
%    Rutgers University, New Brunswick, NJ, USA}
% }

\begin{document}
\maketitle


%\author{Kristof Farkas-Pall, David W Wright, Shunzhou Wan, Peter V Coveney}
%\orcid{}
%\affiliation{%
%  \institution{UCL}
%  \streetaddress{20 Gordon Street}
%  \city{London}
  %\state{}
%  \country{UK}


%\renewcommand\shortauthors{Dakka, J. et al}

\begin{abstract}

The recently demonstrated ability to perform accurate, precise and rapid
computation of ligand binding strengths has driven interest in the use of
molecular simulations for applications in both computer-aided drug design and
patient specific medicine. Simulation protocols based on ensembles of multiple
runs of the same system provide an efficient method for producing robust free
energy estimate, and equally important statistical uncertainties. Variations
in the chemical and biophysical properties of different systems impact the
optimal protocol choice for different proteins and classes of drugs targetting
them. However, these are difficult to determine {\it a priori}, thus requiring
adaptive execution of a simulation workflows. We introduce the High-throughput
Binding Affinity Calculator (HTBAC) to enable the scalable, adaptive and
automated calculation of the binding free energy on high-performance computing
resources. We investigate weak scaling behaviour for screening sixteen drug
candidates concurrenlty using thousands of multi-stage pipelines on more than
32,000 cores. This permits a rapid time-to-solution that is essentially
invariant with respect to the calculation protocol, size of candidate ligands
and number of ensemble simulations. As such, HTBAC advances the state of the
scale and sophistication of binding affinity calculation. HTBAC uses readily
available building blocks to attain both workflow flexibility and performance;
our scaling experiments are performed on the Blue Waters machine at NCSA.

\end{abstract}


% To simply model building and simulation setup we developed an automation tool,
% the Binding Affinity Calculator (BAC). Building on this tool, to support the
% scalable, adaptive and automated calculation of the binding free energy on
% high-performance computing resources,


% \begin{abstract}
% Recent improvements in computational power and algorithm design mean that
% reliably quantifying ligand binding strength differences from molecular
% simulation is becoming a genuine possibility. These developments have led to
% increased interest in the use of molecular simulations for applications in both
% computer-aided drug design and patient specific medicine. We have previously
% demonstrated that simulation protocols based on ensembles of multiple runs of
% the same system provide an efficient method for producing robust free energy
% estimate, and equally important statistical uncertainties. Variations in the
% chemical and biophysical properties of different systems impact the optimal
% protocol choice for different proteins and classes of drugs targetting them.
% This leads to the necessity of automation and middleware solutions that allow
% adaptive execution of a wide range of simulation workflows. To support the
% scalable, adaptive and automated calculation of the binding free energy on
% high-performance computing resources, we introduce the High- throughput Binding
% Affinity Calculator (HTBAC). HTBAC uses a building block approach in order to
% attain both workflow flexibility and performance. We demonstrate close to
% perfect weak scaling to hundreds of concurrent multi-stage binding affinity
% calculation pipelines. This permits a rapid time-to-solution that is
% essentially invariant of the calculation protocol, size of candidate ligands
% and number of ensemble simulations. As such, HTBAC advances the state of the
% art of binding affinity calculations and protocols.
% \end{abstract}


\keywords{binding free energy, binding affinity, drug binding, middleware}

%\maketitle

\section{Introduction}
\input{introduction}

%\section{Related Work}
%\input{related_work}

\section{Related Work}
\label{sec:related-work}
\input{related_work2}


\section{Science Drivers}
\label{sec:science-drivers}
\input{fe_protocols}
\input{brd4}

\section{HTBAC: Design and Implementation}
\input{design_implementation}

\section{Experiments}
\input{experiments}

%\section{Results}
%\input{results}

\section{Discussion and Conclusion}
\input{conclusions}

\subsubsection{Acknowledgements}
%\input{acknowledgements}

{\footnotesize JD acknowledges Dept. of Education Award P200A150217 for her fellowship. PVC
would like to acknowledge the support of the EU H2020 CompBioMed (675451),
EUDAT2020 (654065) and ComPat (671564) projects. We thank Andre Merzky and
other members of the RADICAL team for support. Access to Blue Waters was made
possible by NSF 1713749. The software capabilities were supported by
RADICAL-Cybertools (NSF 1440677).}

\newpage

% ---------------------------------------------------------------------------
% BIBLIOGRAPHY
% -------------------,--------------------------------------------------------
\bibliographystyle{IEEEtran}
\bibliography{rutgers,ucl}

\end{document}
