% Science drivers --------------------------------------------------------------
% Everything related to the chemistry/drug discovery science in this paper. ----

The drug discovery pipeline is continuously evolving and morphing from a fully experimental procedure, to stages of it being replaced by computational methods, like computer aided drug design. In its simplest form, the aim of a drug discovery campaign is to optimize the efficacy of a drug molecule to a \emph{certain} target (usually a protein) as a function of the drug molecule itself. In the \emph{hit to lead}, and \emph{lead optimization} stages of the pipeline this is done by maximizing a quantity called the binding affinity (which is a good proxy for efficacy of the medication at this stage). To find the a maximum we have to probe large number of candidate drug molecules and find the best ones. The basic use case for HTBAC is to enable this type of large scale binding free energy studies of protein-ligand complexes using ensemble simulations. 

\subsection{ESMACS and TIES}

% Intro to binding affinity ----------------------------------------------------

There exists two types of free energy calculation protocols: absolute and 
relative. Absolute free energy methods calculate the binding affinity of a 
\emph{single} drug molecule to a protein, while relative methods calculate the 
\emph{difference} in binding affinity between two (usually similar in 
structure) drug molecules. The computational requirements of the two protocols 
differ sometimes significantly, but deciding which one is more appropriate for 
the task at hand is not trivial, as it is a function of resources, required 
time to convergence, accuracy thresholds etc.

Briefly introduced in the previous section, the two free energy calculation 
protocols, ESMACS and TIES ~\cite{Wan2017brd4, Bhati2017}, are our 
implementations of the absolute and relative methods. Here we provide more 
details of their practical specifications and of the protein ligand systems we 
used to benchmark and refine them in this work.

% Some notes on ensembles and replicas. ----------------------------------------

In both protocols, ensembles of MD simulations are employed to perform 
averaging and to obtain tight control of error bounds in our estimates. In 
addition, the ability to run replica simulations concurrently means that, as 
long as sufficient compute resources are available, turnaround times can be  
significantly reduced compared to the generation of single long trajectories. 
Due to their shared philosophical underpinning both protocols share similar 
middleware requirements.

% First a quick intro to NAMD, then generic simulation protocol description. ---

The current implementation of both TIES and ESMACS uses the the NAMD 
package~\cite{Phillips2005} to conduct the simulations. Conceptually, each 
replica simulation consists of three stages: minimization, equilibration and 
production MD. In practice the equilibration phase is broken into multiple 
steps to ensure that the size of the simulation box does not alter too much 
over the simulation. During these steps positional constraints are gradually 
released from the structure and the system is heated to a physiologically 
realistic temperature. Whilst both protocols share a common workflow for 
individual replicas, the make up of the ensemble is different.

% TIES and ESMACS intro --------------------------------------------------------

TIES is based on rigorous, but computationally expensive, calculations of 
relative free energies (i.e. results provide a comparison between two drugs). 
ESMACS, in contrast, provides absolute binding free energies at low 
computational cost, but to achieve this coarse grains many of the details of 
the system being studied. The simplifications employed by ESMACS can reduce its 
effectiveness in some systems. In the case of ESMACS, an ensemble consists of a 
set of 25 identical simulations differing only in the initial velocities 
assigned to each atom. In TIES the ensemble also contains the set of $\lambda$ 
windows along the replicas. Upon completion of the MD simulation, free energy 
computations are performed on a per replica basis and then globally across all 
ensembles. 

%~\cite{amber14, Case2005, MillerIII2012}.

% TIES more details ------------------------------------------------------------



TIES is a so called ``alchemical'' method in which the fact that the potential 
used to describe the system is user definable to transform one system into 
another. This allows for the calculation of free energy differences between 
the two systems, such as those induced by an alteration to a candidate drug.
Typically a transformation parameter (called $\lambda$) is defined such 
that as it increases from zero to one the system description transforms from 
containing an initial drug to a target compound via a series of hybrid states.
Sampling along $\lambda$ is then required in order to compute the difference 
in binding free energy. Consequently, whereas in ESMACS all replicas sample 
using the same system description, in TIES sub-ensembles are executed at 
different points along $\lambda$. The change in free energy associated 
with the transformation is calculated by numerically integrating 
(via adaptive quadrature) the values of the $\partial U/\partial\lambda$ 
across the full set of $\lambda$ windows simulated. 

Obtaining accurate and precise results from TIES requires that the $\lambda$ 
windows correctly capture the changes of $\partial U/\partial\lambda$ over 
the transformation. This behavior is highly sensitive to
the chemical details of the compounds being studied and varies considerably 
between systems. Typically, windows are evenly spaced between 0 and 1 with 
the spacing between them set before execution at a distance determined by the 
simulator to be sufficient for a wide range of systems. Typically, a TIES 
ensemble 65 replicas evenly distributed between 13 $\lambda$ windows
In order to obtain a meaningful TIES result it is necessary to not only 
simulate the drug pair in the protein but also in an aqueous environment, 
adding a further 65 replicas albeit it using a smaller system at lower 
computational cost.

% Analysis ---------------------------------------------------------------------

Following simulation (and in the case of ESMACS free energy analysis steps) 
both protocols require the execution of short serial steps to provide summary 
statistics. Both protocols can are highly customizable, for example the number 
of simulation replicas in the ensemble and the lengths of their runs can be 
varied. More generally though, there may be cases where it is important to 
increase the sampling of phase space possibly through expanding the ensemble 
or in TIES changing the distribution of $\lambda$ windows.
In addition, the ESMACS protocol can also be extended to account for 
adaptation energies involved in altering the conformation of the protein 
or ligand during binding through the use of separate component simulations.

%Adaptive quadrature is a numerical integration algorithm that is designed to
%minimize the number of function evaluations while still achieving a given error
%tolerance when estimating the integral. 
%We use this algorithm to minimize the
%number of full simulations that are required to be run, while still keeping the
%high accuracy of the protocol.

\subsection{The Value of Adaptivity}

Both ESMACS and TIES have been successfully used to predict binding affinities
quickly and accurately. Nonetheless, they are very expensive computationally,
and optimizing the execution time while still improving the accuracy is
desirable. Given the very large number of drug candidates, it is imperative to
gain maximum insight into potential candidate compounds using time and
resources efficiently. This provides one clear motivation for the use of
adaptive methods which minimize the compute time used whilst producing binding
free energy estimates meeting pre-defined quality criteria (such as
convergence or statistical uncertainty below a given threshold).
  
% Drug discovery programmes usually have limited resources, and require a
% resource budget carefully looking to gain maximum insight into potential
% candidate compounds. 

A second driver for adaptivity is that such algorithmic methods  will
typically involve compounds with a wide range of chemical properties which can
impact not only the time to convergence, but the type of sampling required to
gain accurate results. In general, there is no way to know before running
calculations exactly which setup of calculation is required for a particular
system. For example in TIES, the number (or the exact location) of the
$\lambda$ windows that will most impact the calculation are not known
\textit{a priori}, and change between physical systems (drugs). As multiple
simulations must be run for each window, sampling with a very high frequency
is expensive and impractical. Furthermore, adaptive placement of $\lambda$
windows is likely to better capture the shape of the 
$\partial U/\partial\lambda$ curve, leading to more accurate and precise 
results for a given computational cost.

On occasion, alchemical methods may be very slow to converge; in such
circumstances use of another method, such as ESMACS, may be the best option.
This means that the most effective way to gain accurate and precise free
energy results on industrially or clinically relevant timescales is to be able
to adapt both sampling (intra-protocol) and even the type of calculation 
(inter-protocol) used at run time. With potentially thousands of simulations, 
often employing multiple analysis methodologies, this provides the most 
effective way to utilize these techniques and resources at scale.

