%\documentclass[conference,final]{IEEEtran}
\documentclass[conference]{IEEEtran}

\input{head}
\input{include}

\begin{document}

\title{Concurrent and Adaptive Extreme Scale Binding Free Energy
Calculations}

\author{
	\IEEEauthorblockN{
		Jumana Dakka\IEEEauthorrefmark{1}, 
		Kristof Farkas-Pall\IEEEauthorrefmark{2}, 
		Matteo Turilli\IEEEauthorrefmark{1}, \\ 
		David W. Wright\IEEEauthorrefmark{2},  
		Peter V. Coveney\IEEEauthorrefmark{2}, 
		Shantenu Jha\IEEEauthorrefmark{1}\IEEEauthorrefmark{3}
	}
	\IEEEauthorblockA{
		\IEEEauthorrefmark{1}Rutgers University, Piscataway, NJ, USA
	}
	\IEEEauthorblockA{
		\IEEEauthorrefmark{2}Center for Computational Sciences, University College London, UK
	}
	\IEEEauthorblockA{
		\IEEEauthorrefmark{3}Brookhaven National Laboratory, Upton, NY, USA
	}
}

\maketitle

\begin{abstract}
The efficacy of drug treatments depends on how tightly small molecules bind to
their target proteins. The rapid and accurate quantification of the strength
of these interactions (as measured by `binding affinity’) is a grand challenge
of computational chemistry, surmounting which could revolutionize drug design
and provide the platform for patient-specific medicine. Recent evidence
suggests that molecular dynamics (MD) can achieve useful predictive accuracy
($\leq$ 1 kcal/mol). For this predictive accuracy to impact clinical decision
making, binding free energy computational campaigns must provide results
rapidly and without loss of accuracy. This demands advances in algorithms,
scalable software systems, and efficient utilization of supercomputing
resources. We introduce a framework called HTBAC, designed to support accurate
and scalable drug binding affinity calculations, while marshaling large
simulation campaigns. We show that HTBAC supports the specification and
execution of free-energy protocols at scale. This paper makes three main
contributions: (1) shows the importance of adaptive execution for 
ensemble-based free energy protocols to improve binding affinity accuracy; (2)
presents and characterizes HTBAC -- a software system that enables the scalable
and adaptive execution of binding affinity protocols at scale; and (3) for a
widely used free-energy protocol (TIES), shows improvements in the accuracy of
simulations for a fixed amount of resource, or reduced resource consumption
for a fixed accuracy as a consequence of adaptive execution.

\end{abstract}


% ---------------------------------------------------------------------------
% INTRODUCTION
% ---------------------------------------------------------------------------
\section{Introduction}\label{sec:intro}
\input{introduction}


% ---------------------------------------------------------------------------
% S2
% ---------------------------------------------------------------------------
\section{Motivation}\label{sec:related-work}
\input{related_work2}


% ---------------------------------------------------------------------------
% S3
% ---------------------------------------------------------------------------
\section{Science Drivers}\label{sec:science-drivers}
\input{fe_protocols}
\input{physical_systems}


% ---------------------------------------------------------------------------
% S4
% ---------------------------------------------------------------------------
\section{High-Throughput Binding Affinity Calculator
(HTBAC)}\label{sec:htbac}
\input{design_implementation}


% ---------------------------------------------------------------------------
% S6
% ---------------------------------------------------------------------------
\section{Experiments}\label{sec:experiments}
\input{experiments}


% ---------------------------------------------------------------------------
% S7
% ---------------------------------------------------------------------------
\section{Discussion and Conclusion}\label{sec:discussion}
\input{conclusions}


% ---------------------------------------------------------------------------
% Software and Data
% ---------------------------------------------------------------------------
\section*{Software and Data}
All experimental data can be found at
\url{https://github.com/radical-experiments/htbac-escience-18}. HTBAC (MIT
license) can be found at: \url{https://github.com/radical-cybertools/htbac}

% ---------------------------------------------------------------------------
% Acknowledgments
% ---------------------------------------------------------------------------
\section*{Acknowledgments}
JD acknowledges Dept. of Education Award P200A150217 for her
fellowship. PVC acknowledges the support of the EU H2020 CompBioMed (675451),
EUDAT2020 (654065) and ComPat (671564) projects. Access to Blue Waters was
made possible by NSF 1713749. The software capabilities were supported by
RADICAL-Cybertools (NSF 1440677). We thank Vivek Balasubramanian and Andre
Merzky for support with RADICAL-Cybertools.


% ---------------------------------------------------------------------------
% BIBLIOGRAPHY
% ---------------------------------------------------------------------------
\bibliographystyle{IEEEtran}
\bibliography{rutgers,ucl}

\end{document}
