\subsection{Physical system description}

The scientific and or computational improvements require validation across a number of protein ligand complexes. We selected 4 proteins and 8 ligands or ligand pairs to run the adaptive relative free energy calculations. The systems were previously used for non-adaptive simulations by \cite{} et. al., offering a baseline comparison against our results. 

The selected proteins are the Protein tyrosine phosphatase 1B (PTP1B), the Induced myeloid leukemia cell differentiation protein (MC1), tyrosine kinase 2 (TYK2) and the bromodomain-containing protein 4 (BRD4). The ligand complexed with the protein fall into two categories. Four of the eight ligands are alchemical transformations from one to another (used in the TIES workflow), the rest being single ligands suitable for absolute free energy calculations, in our case for ESMACS.

Preparation and setup of the simulations were implemented using our automated tool, BAC~\cite{Sadiq2008}. This process includes parametrization of the compounds, solvation of the complexes, electrostatic neutralization of the systems by adding counterions and generation of configurations files for the simulations. The AMBER ff99SB-ILDN \cite{Lindorff-Larsen2010} force field was used for the proteins, and TIP3P was used for water molecules. Compound parameters were produced using the general AMBER force field (GAFF)~\cite{Wang2004} with Gaussian 03~\cite{Frisch} to optimize compound geometries and to determine electrostatic potentials at the Hartree–Fock level with 6-31G** basis functions. The restrained electrostatic potential (RESP) module in the AMBER package~\cite{Case2005} was used to calculate the partial atomic charges for the compounds. All systems were solvated in orthorhombic water boxes with a minimum extension from the protein of 14 \AA\ (the resulting systems contain approximately 40 thousand atoms).

% \subsection{Target protein: BRD4}
% 
% Bromodomain-containing proteins, and in particular the four members of the BET
% (bromodomain and extra terminal domain) family, are currently a major focus of
% research in the pharmaceutical industry. Small molecule inhibitors of these
% proteins have shown promising preclinical efficacy in pathologies ranging from
% cancer to inflammation. Indeed, several compounds are progressing through
% early stage clinical trials and are showing exciting early
% results~\cite{Theodoulou2016}. One of the most extensively studied targets in
% this family is the first bromodomain of bromodomain-containing protein 4
% (BRD4-BD1) for which extensive crystallographic and ligand binding data are
% available~\cite{Bamborough2012}.
% 
% We have previously investigated a congeneric series of ligands binding to
% BRD4-BD1 (we shall from now on refer to this are simply BRD4) using both
% ESMACS and TIES. This was performed in the context of a blind test of the
% protocols in collaboration with GlaxoSmithKline~\cite{Wan2017brd4}. The goal
% was to benchmark the ESMACS and TIES protocols in a realistic drug discovery
% scenario. In the original study, we investigated chemical structures of 16
% ligands based on a single tetrahydroquinoline (THQ) scaffold.
% % ~\cite{Gosmini2014}. 
% Here we focus on the first seven of these ligands to test
% and refine the protocols used and the way in which they were executed. The
% results of our previous work provide a benchmark of both accuracy and
% statistical uncertainty to which we can compare our new results.
% 
% Initial coordinates for the protein-ligand system were taken from the X-ray
% crystal structure PDB ID: 4BJX.
% % ~\cite{Wyce2013}. 
% This structure contains a
% ligand based on the THQ template and other ligands were aligned with this
% common scaffold. 
